---
title: "GSS challange"
author: "Leilani Guerrero"
date: "Summer 2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$: Attitude towards capital punishment is not related to race.

$H_A$: Attitude towards capital punishment is related to race.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
Both race and cappun are categorical variables so the analysis technique is CAT~CAT, plot is a box plot, summary is a cross table, inference is chi-squared test.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


According to the chart the race "white" has an overall greater percentage (73%) of being in favor of cappun, with race "other" holding (55%), and race "black" last with 45% in favor of cappun. However, race "black" has an overall greater percentage of 53% in opposed to cappun, with other second with 43% and race "white" in last with 32% in opposed to cappun.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with More data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```

According to the number chart, out of 100 race "white" 67.98 are in favor of cappun and 32.02 are opposed to cappun. Out of 100 race "black" 46.29 are in favor of cappun and 53.71 are opposed to cappun. Out of 100 race "other" 55.87 are in favor of cappun and 44.13 are opposed to cappun. In addition, out of the 100 races in favor of cappun; 78.34 are of "white" race, 11.7 are of "black" race, and 9.96 are other. Out of the 100 of races that are opposed to cappun; 63.24 are white, 23.27 are black, and 13.49 are other.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

The p-value for the chi-squared test was 0.00000000000001027, because the p-value is less then 0.05 we reject the null hypothesis.


```{r}
fisherXY(mydata$cappun,mydata$race)
```
 
 For the first compassion, White v Black, 1085 of white are in favor of cappun and 511 of white are opposed of cappun, 162 of black are in favor of cappun and 188 of black are opposed to cappun. because the p-value of white v black is 0.0000000000000575 which is less than 0.05 we reject the null hypothsis. 
 For the second compassion, White v Other, 1085 of white are in favor of cappun and 511 of white are opposed of cappun, 138 of other are in favor of cappun and 109 of other are opposed to cappun. because the p-value of white v black is 0.0002819 which is less than 0.05 we reject the null hypothsis.
 For the last compassion, Other v Black, 138 of other are in favor of cappun and 109 of other are opposed of cappun, 162 of black are in favor of cappun and 188 of black are opposed to cappun. because the p-value of white v black is 0.02479 which is less than 0.05 we reject the null hypothsis.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
The null hypothesis was Attitude towards capital punishment is not related to race., while the alternative Attitude towards capital punishment is related to race. However, since the p-value for the chi-squared test was 0.00000000000001027 we reject the null hypothesis because the P-value is less than 0.05. 